Theories of deliberated choice

### 2023-10-10 simple.tex
Equiv fals proven

Given an observation $\theta$, let $\theta^{-1}(P) \subseteq I × \allhist$ denote the pairs $(i, \hist)$ in $\theta$.
Let $\incr(\allhist) = \cup_{N \subseteq \N} \set{A \in (\allhist)^N \suchthat \card{A_i} = i \land i < j \implies A_i \subsetneq A_j} \subseteq \powerset{\allhist}$ denote the possibly empty embedded sequences of paths.
An element $\theta^{-1}(P)$ can also be seen as a (finite) function associating to each $i \in I$ an element of $\incr(\allhist)$, thus, a possibly empty embedded sequence of elements in $\allhist$.
Therefore, we let $\incr(\allhist)^I = \set{\theta^{-1}(P) \suchthat \theta \in \Theta}$ denote the set of possible observations restricted to $I × \allhist$, the elements of which being called \emph{a priori observation}.
Note that $\incr(\allhist)^I$ is downwards closed, meaning that $\forall O \in \incr(\allhist)^I: O' \subseteq O \implies O' \in \incr(\allhist)^I$.
Let also $\max\incr(\allhist) = \set{A \in \incr(\allhist) \suchthat \nexists A' \in \incr(\allhist) \suchthat A \subset A'} = \set{A \in \incr(\allhist) \cap (\allhist)^\N}$ denote the maximal elements of $\incr(\allhist)$, in the sense of containing sequences that are not a subsequence of another sequence; equivalently, containing those sequences whose sequence of cardinals of their elements equals $\N$.
Let a \emph{simple a priori observation} be defined as an element $O \in (\allhist)^I$.
Given a simple a priori observation $O: I \mapsto \allhist$, we can define the a priori observation $O_O = \set{(i, \hist) \in I × \allhist \suchthat \forall i \in I: \hist \subseteq O(i)}$ as the completion of $O$ to all sub-paths in $O$. 
Given $O$ and ${\allleadstog}$, it follows that $S_{\restr{\allleadstog}{O_O}}$ denotes the “leads to” relations that complete $O_O$ using $\allleadstog$.

@@

    To obtain non fals, need to form a relation whose all decompositions into incr associations are in $C$, but itself not in $C$.
  \end{remark}
  \begin{conjecture}
    Given a claim $C$, consider the subsets of $(\cup_{A \in \max \incr(\allhist)}P^A)^I \cap C^\subseteq$ whose union, considered as a relation on $(I × \allhist) × P$, is total and right-unique (that is, the subsets that form a new relation $\allleadstog$).
    $C$ is falsifiable iff these new elements are necessarily in $C$.
  \end{conjecture}
  \begin{proof}
    This seems false. (Consider $I = \set{i_1, i_2}$ and $\args = \set{a, b}$ and $\psi$ is ${\ileadsto[i_1]} = {\ileadsto[i_2]}$ and $S = \set{R_1, R_2, …}$ with $R_1$ are all sequences starting with $a$ sent to $p$ and $R_2$ are all sequences starting with $b$ sent to $p' ≠ p$.)
    % Consider $S \subseteq (\cup_{A \in \max \incr(\allhist)}P^A)^I \cap C^\subseteq$ (thus each $R \in S$ is a relation $R \subseteq (I × \allhist) × P$). Assume that $\cup S$ is total and right-unique and is not in $C$.
    % Let us show that $C$ is not falsifiable.
    % Define ${\allleadstog} = \cup S$, pick any $\theta \in \Theta, \theta \subseteq {\allleadstog}$ and let us show that $\theta \in C^\subseteq$.
    % Let us show that $\exists R \in S \suchthat \theta \subseteq R$.
    % Pick any triple $t \in \theta$. Thus, $t \in \cup S$. Thus, for some $R \in S$, $t \in R$. And $R \in C^\subseteq$.
    % Pick any two triples $t, t' \in \theta$. Wlog, $\hist \subseteq \hist'$.
  \end{proof}
  \begin{conjecture}
    Given a claim $C$, consider the sets $S \subseteq C^\subseteq$ such that $\set{s^{-1}(P) \suchthat s \in S} = \incr(\allhist)^I$ and such that $\cup S$, considered as a relation on $(I × \allhist) × P$, is right-unique (that is, $\cup S$ forms a new relation $\allleadstog$).
    $C$ is falsifiable iff all such sets $S$ are in $C$.
  \end{conjecture}
  \begin{proof}
    Consider such $S$.
    Assume that $\cup S$ is right-unique and is not in $C$.
    Let us show that $C$ is not falsifiable.
    Define ${\allleadstog} = \cup S$, pick any $\theta \in \Theta, \theta \subseteq {\allleadstog}$ and let us show that $\theta \in C^\subseteq$.
    Define $O \in (\incr(\allhist))^I \suchthat \theta^{-1}(P) \subseteq O$. Thus, $O \in \set{s^{-1}(P) \suchthat s \in S}$. Pick an $s \in S \suchthat s^{-1}(P) = O$. We have $s \in C^\subseteq$ and $\theta^{-1}(P) \subseteq s^{-1}(P)$.
    Also, $\theta \subseteq \cup S$, and $\cup S$ is right-unique, so $\theta$ coincides with $s$ on its domain. It follows that $\theta \subseteq s$. As $C^\subseteq$ is downwards-closed, this concludes.

    Now, Assume that $C$ is not falsifiable and pick any ${\allleadstog} \notin C \suchthat \forall \theta \in \Theta, \theta \subseteq {\allleadstog}: \theta \in C^\subseteq$. Let us exhibit a set $S$ such that $\cup S = {\allleadstog}$.
    Pick any $O \in (\incr(\allhist))^I$ and complete it with values from $\allleadstog$ to obtain a $c$. Now, we know that $C^\subseteq$ contains all finite subsequences of $c$, but it does not follow that it contains the infinite sequence $O$, does it?
    If $C ≠ \emptyset$, an element $c'$ has the same domain as $c$. We know that $c_1 \subseteq c_2 \subseteq …$, where each is a finite subset, is extensible to some function in $C$… ???
  \end{proof}

>>>>

Given an observation $\theta$, let $\theta^{-1}(P) \subseteq I × \allhist$ denote the pairs $(i, \hist)$ in $\theta$.
Let $\incr(\allhist) = \cup_{N \subseteq \N} \set{A \in (\allhist)^N \suchthat [\card{A_i} = i] \land [i < j \implies A_i \subsetneq A_j]} \subseteq \powerset{\allhist}$ denote the possibly empty embedded sequences of paths.
An element $\theta^{-1}(P)$ can also be seen as a (finite) function associating to each $i \in I$ an element of $\incr(\allhist)$, thus, a possibly empty embedded sequence of elements in $\allhist$.
Therefore, we let $\incr(\allhist)^I = \set{\theta^{-1}(P) \suchthat \theta \in \Theta}$ denote the set of possible observations restricted to $I × \allhist$, the elements of which being called \emph{a priori observation}.
Note that $\incr(\allhist)^I$ is downwards closed, meaning that $\forall O \in \incr(\allhist)^I: O' \subseteq O \implies O' \in \incr(\allhist)^I$.
Let also $\max\incr(\allhist) = \set{A \in \incr(\allhist) \suchthat \nexists A' \in \incr(\allhist) \suchthat A \subset A'} = \set{A \in \incr(\allhist) \cap (\allhist)^\N}$ denote the maximal elements of $\incr(\allhist)$, in the sense of containing sequences that are not a subsequence of another sequence; equivalently, containing those sequences whose sequence of cardinals of their elements equals $\N$.

Let the \emph{simple a priori observations} be defined as $(\allhist)^I$.
Given a simple a priori observation $O: I \mapsto \allhist$, we can define the a priori observation $O_O = \set{(i, \hist) \in I × \allhist \suchthat \forall i \in I: \hist \subseteq O(i)}$ as the completion of $O$ to all sub-paths in $O$.
Given $O$ and ${\allleadstog}$, it follows that $S_{\restr{\allleadstog}{O_O}}$ denotes the “leads to” relations that complete $O_O$ using $\allleadstog$.

@@

  To obtain non fals, need to form a relation whose all decompositions into incr associations are in $C$, but itself not in $C$.
\end{remark}
\begin{conjecture}
  Given a claim $C$, consider the subsets of $(\cup_{A \in \max \incr(\allhist)}P^A)^I \cap C^\subseteq$ whose union, considered as a relation on $(I × \allhist) × P$, is total and right-unique (that is, the subsets that form a new relation $\allleadstog$).
  $C$ is falsifiable iff these new elements are necessarily in $C$.
\end{conjecture}
\begin{proof}
  This seems false. (Consider $I = \set{i_1, i_2}$ and $\args = \set{a, b}$ and $\psi$ is ${\ileadsto[i_1]} = {\ileadsto[i_2]}$ and $S = \set{R_1, R_2, …}$ with $R_1$ are all sequences starting with $a$ sent to $p$ and $R_2$ are all sequences starting with $b$ sent to $p' ≠ p$.)
  % Consider $S \subseteq (\cup_{A \in \max \incr(\allhist)}P^A)^I \cap C^\subseteq$ (thus each $R \in S$ is a relation $R \subseteq (I × \allhist) × P$). Assume that $\cup S$ is total and right-unique and is not in $C$.
  % Let us show that $C$ is not falsifiable.
  % Define ${\allleadstog} = \cup S$, pick any $\theta \in \Theta, \theta \subseteq {\allleadstog}$ and let us show that $\theta \in C^\subseteq$.
  % Let us show that $\exists R \in S \suchthat \theta \subseteq R$.
  % Pick any triple $t \in \theta$. Thus, $t \in \cup S$. Thus, for some $R \in S$, $t \in R$. And $R \in C^\subseteq$.
  % Pick any two triples $t, t' \in \theta$. Wlog, $\hist \subseteq \hist'$.
\end{proof}
\begin{theorem}
  Given a claim $C$, consider the sets $S \subseteq C^\subseteq$ such that $\set{\theta^{-1}(P) \suchthat \theta \in S} = \incr(\allhist)^I$ and such that $\cup S$, considered as a relation on $(I × \allhist) × P$, is right-unique (that is, $\cup S$ forms a new relation $\allleadstog$).
  $C$ is falsifiable iff all such sets $S$ satisfy $\cup S \in C$.
\end{theorem}
\begin{proof}
  Consider such $S$.
  Assume that $\cup S$ is not in $C$.
  Let us show that $C$ is not falsifiable.
  Define ${\allleadstog} = \cup S$, pick any $\theta \in \Theta, \theta \subseteq {\allleadstog}$ and let us show that $\theta \in C^\subseteq$.
  Define $O \in (\incr(\allhist))^I \suchthat \theta^{-1}(P) \subseteq O$. Thus, $O \in \set{\theta^{-1}(P) \suchthat \theta \in S}$. Pick an element $\theta' \in S \suchthat (\theta')^{-1}(P) = O$. We have $\theta' \in C^\subseteq$ and $\theta^{-1}(P) \subseteq (\theta')^{-1}(P)$.
  Also, $\theta \subseteq \cup S$, and $\cup S$ is right-unique, so $\theta$ coincides with $\theta'$ on its domain (meaning that $\theta = \restr{\theta'}{\theta^{-1}(P)}$). It follows that $\theta \subseteq \theta'$. As $C^\subseteq$ is downwards-closed, this concludes.

  Now, assume that $C$ is not falsifiable and pick any ${\allleadstog} \notin C \suchthat \forall \theta \in \Theta, \theta \subseteq {\allleadstog}: \theta \in C^\subseteq$.
  Given any $O \in (\incr(\allhist))^I$, we have $\restr{\allleadstog}{O} \in C^\subseteq$.
  Define $S = \set{\restr{\allleadstog}{O} \suchthat O \in (\incr(\allhist))^I}$.
  The set $S$ satisfies the required conditions and $\cup S = {\allleadstog}$, thus, $\cup S \notin C$.
\end{proof}
Here is another way (perhaps more illuminating) of stating the same thing.
Any relation $\allleadstog$ is equivalent to $\cup \set{\restr{\allleadstog}{O} \suchthat O \in (\incr(\allhist))^I}$.
Conversely, any set $S$ such that $\set{\theta^{-1}(P) \suchthat \theta \in S} = \incr(\allhist)^I$ and $\cup S$ is right-unique corresponds to a unique relation $\cup S = {\allleadstog}$.
Thus, there is a bijection between the set of such sets $S$ and the set of leads to relations.
We can then rephrase the above theorem as follows.
A claim $C$ is falsifiable iff $\forall S \in \mathcal{S}: S \subseteq C^\subseteq \implies \cup S \in C$.
Equivalently, as follows.
\begin{theorem}
  A claim $C$ is falsifiable iff $\forall {\allleadstog} \in P^{I × \allhist}: \set{\restr{\allleadstog}{O} \suchthat O \in (\incr(\allhist))^I} \subseteq C^\subseteq \implies {\allleadstog} \in C$.
\end{theorem}
\begin{proof}
  By definition, $C$ is not falsifiable iff $\exists {\allleadstog} \notin C \suchthat \forall θ \subseteq {\allleadstog}: S_θ \cap C \neq \emptyset$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \forall θ \subseteq {\allleadstog}: θ \in C^\subseteq$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \set{θ \subseteq {\allleadstog}} \subseteq C^\subseteq$.

  The theorem holds because $\forall {\allleadstog} \in P^{I × \allhist}: \set{\restr{\allleadstog}{O} \suchthat O \in (\incr(\allhist))^I} = \set{θ \subseteq {\allleadstog}}$.
\end{proof}
